---
title: "string_practice"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
alarm_report<-tibble::tribble(
         ~date,           ~building,                                               ~alarm_message,
  "2020-03-14", "Engineering-North", "10:02am -- HVAC system down, facilities management alerted",
  "2020-03-15",         "Bren Hall",                    "8:24am -- Elevator North out of service",
  "2020-04-10", "Engineering-South",     "12:41am -- Fire alarm, UCSB fire responded and cleared",
  "2020-04-18",        "Engr-North",     "9:58pm -- Campus point emergency siren, UCPD responded"
  )

view(alarm_report)
```


Practice working with strings by writing code to update alarm_report as follows (these can be separate, or all as part of a piped sequence):

Replace the “Engr” with “Engineering” in the building column
Separate the building column into two separate columns, building and wing, separated at the dash
Only keep observations with the word “responded” in the alarm_message column
Separate the message time from the rest of the message by separating at --
Convert the date column to a Date class using lubridate

```{r}
#Replace the “Engr” with “Engineering” in the building column
update_alarm_report1<- alarm_report %>% 
  mutate(building = str_replace(building,
                            pattern = "Engr-",
                            replacement = "Engineering-"))

#Separate the building column into two separate columns, building and wing, separated at the dash
update_alarm_report2 <- update_alarm_report1 %>%
  separate(building, into = c("building", "wing"), sep = "-")

#Only keep observations with the word “responded” in the alarm_message column
update_alarm_report3<-update_alarm_report2%>%
  filter(str_detect(alarm_message,"responded"))

#Separate the message time from the rest of the message by separating at --
update_alarm_report4 <- update_alarm_report2 %>%
  separate(alarm_message, into = c("message time", "message"), sep = "--")

#Convert the date column to a Date class using lubridate
update_alarm_report5 <- update_alarm_report4 %>%
  mutate(date = ymd(date))

class(update_alarm_report5$date)


```

